---
title: "Bella Beat Case Study"
author: "Lusui"
date: "12/27/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# BELLA BEAT CASE STUDY
## Business task:Identify trends in smart device usage for;
(a) Applying said trends to Bella beat users
(b) Influencing the Bella Beat Marketing strategy

### Setting up environment
Notes:The following packages were used(description shown)
```{r setup, setting up environment}
library(tidyverse)#For data cleaning and organizing
library(ggplot2)#For plotting
library(knitr)#For use with rmarkdown oin presentation
library(lubridate)#For sdate use and conversion
library(readr)#For importing data
library(dplyr)#For making manipulations
library(rmarkdown)#For creating a sharable document
library(janitor)#For easing data cleaning
library(skimr)#For easy summary
```


### IMPORT THE DATA(PREPARE)
Notes:Three data sets were imported for this analysis
```{r data import}
#used only two data sets that covered 24 hours(initially)
#added weight data set to get context
daily_activity <- read_csv("C:\\Users\\Allan\\OneDrive - CM Advocates, LLP\\Documents\\R\\Fitabase Data 4.12.16-5.12.16\\dailyActivity_merged.csv")
daily_sleep <- read_csv("C:\\Users\\Allan\\OneDrive - CM Advocates, LLP\\Documents\\R\\Fitabase Data 4.12.16-5.12.16\\sleepDay_merged.csv")
weight_log <- read_csv("C:\\Users\\Allan\\OneDrive - CM Advocates, LLP\\Documents\\R\\Fitabase Data 4.12.16-5.12.16\\weightLogInfo_merged.csv")
```

### PREVIEW THE DATA AND FIND THE STRUCTURE
Notes: Just the first 6 rows & then look at the structure
```{r preview}
head(daily_activity)
head(daily_sleep)
head(weight_log)
#Structure
str(daily_activity)
str(daily_sleep)
str(weight_log)
```

## CLEANING(PROCESS)
### CHECK FOR ANY DUPLICATES AND MISSING VALUES
Notes: Check is done by different simple functions
```{r cleaning}
#unique users check
n_distinct(daily_activity$Id)
n_distinct(daily_sleep$Id)
n_distinct(weight_log)
#check for any na values
sum(is.na(daily_activity))
sum(is.na(daily_sleep))
sum(is.na(weight_log))#65 na.for the fat column
#check for duplicates
sum(duplicated(daily_activity))
sum(duplicated(daily_sleep))
sum(duplicated(weight_log))
#Conclusuion:there are no missing values in the data set but there are duplicated values in the sleep data set
```

### REMOVE DUPLICATES
Notes: We also remove NAs as a precautionary measure
       On the weight dataset we removed  the fat dataset owing to the hge na numbers on this          column
```{r remove duplicates}
daily_sleep_clean <- daily_sleep%>%
  distinct()%>%
  drop_na()
n_distinct(daily_sleep_clean)
n_distinct(daily_activity)
#dataset 2 cleaned

weight_log_clean <- weight_log%>%
  distinct()%>%
  select(-Fat)
View(weight_log_clean)
#dataset 3 cleaned

#Cleaning dataset 1
clean_names(daily_activity)
clean_names(daily_sleep_clean)
```

### Change the date format
Notes: Change the datatype of the data column, convert format to yyyy-mm-dd and rename it "date"
```{r Date Formatting}
#dataset 1
daily_activity <- daily_activity %>%
  rename(Date = ActivityDate) %>%
  mutate(Date = as_date(Date, format = "%m/%d/%Y"))

#dataset 2
daily_sleep_clean <- daily_sleep_clean %>%
  rename(Date = SleepDay) %>%
  mutate(Date = as_date(Date,format ="%m/%d/%Y %I:%M:%S %p" , tz=Sys.timezone()))

#dataset 3
weight_log_clean <- weight_log_clean %>%
  mutate(Date = as_date(Date,format ="%m/%d/%Y %I:%M:%S %p" , tz=Sys.timezone()))
#confirmation
head(daily_activity)
head(daily_sleep_clean)
head(weight_log_clean)
```

## ANALYZING OUR DATA
### Daily activity data
Notes:We Get summaries of our data
```{r summarize daily activity}
daily_activity%>%
  select(TotalSteps,TotalDistance,SedentaryMinutes,LightlyActiveMinutes,
         FairlyActiveMinutes,VeryActiveMinutes,Calories)%>%
  summary()
#NOTABLE STATISTICS
#average daily steps-7,638 < recommended 10,000(bad)
#average distance covered daily-5.49 miles
#Sedentary Minutes - 991.2 - can cause health problems(3/4 a day)
#Lightly Active Minutes - 192.8
#Fairly ightly Active Minutes- 13.56
#Very Active Minutes - 21.6 > recommended 9.7 mins(good)
#Calories - 2304-(average person burns 1800 a day...to lose weight we need 3500 a day)
#On tarck to lose weight but generally slower
```

### Daily sleep data
```{r summarize daily sleep dataset}
daily_sleep_clean%>%
  select(TotalSleepRecords,TotalMinutesAsleep,TotalTimeInBed)%>%
  summary()
#NOTABLE STATISTICS
#AVG Total Minutes Asleep - 419.2(6.986667 hours)(good-recommended 6-8 hrs)
#AVG Total Time In Bed - 458.5(7.641667hours)
#According to Health Central, people should not spend more than 1 hour in bed awake.
#This is to prevent a mental link being formed between being awake and being in bed, which can lead to insomnia.
```

### Weight dataset
```{r summarize the weight dataset}
weight_log_clean%>%
  select(WeightKg,BMI)%>%
  summary()

#NOTABLE STATISTICS
#AVG BMI - 25.19(overweight)...does not generally indicate body fatness
```

## PLOTS (Share phase)
### DAILY ACTIVITY DATSET

### Total steps vs sedentary minutes
Notes: there is a negative linear relation
```{r Total steps vs sedentary minutes}
#Total steps vs sedentary minutes
summary(daily_activity)
ggplot(daily_activity, aes(x=TotalSteps, y =SedentaryMinutes,color=Calories))+
  geom_point()+geom_smooth(method = lm)+
  labs(title= "Total steps vs sedentary minutes")+
  annotate("text",x=22000,y=350,label="negative correlation", color="red",fontface="bold")
```

### Calories vs very active minutes
Notes: positive linear relation
```{r Calories vs very active minutes}
ggplot(daily_activity, aes(x=VeryActiveMinutes, y =Calories,color=Calories))+
  geom_point()+geom_smooth(method = lm)+
  labs(title= "Very active minutes vs Calories")+
  annotate("text",x=120,y=1000,label="positive correlation", color="green",fontface="bold")
```

### Steps vs Calories
Notes: Positive Linear Relation
```{r Steps vs Calories}
#Steps vs Calories
ggplot(daily_activity, aes(x=TotalSteps, y =Calories,color=Calories))+
  geom_point()+geom_smooth(method = lm)+
  labs(title= "Steps vs Calories")+
  annotate("text",x=30000,y=1000,label="positive correlation", color="green",fontface="bold")
```

## DAILY SLEEP
### Total Minutes Asleep vs TotalTime In Bed
Notes: Positive Linear Relation
```{r minutes asleep vs Total time in bed}
ggplot(daily_sleep_clean, aes(x =TotalMinutesAsleep, y =TotalTimeInBed,))+
  geom_point()+geom_smooth(method = lm)+
  labs(title= "Total Minutes Asleep vs TotalTime In Bed")+
  annotate("text",x=500,y=230,label="positive correlation", color="green",fontface="bold")

```

## WEIGHT LOG
### Weight vs BMI
Notes: Positive Linear Relation
```{r  weigtht vs BMI}
ggplot(weight_log_clean, aes(x=WeightKg, y =BMI))+
  geom_point()+geom_smooth(method = lm)+
  labs(title = "Weight vs BMI")+
  annotate("text",x=70,y=38,label="positive correlation", color="green",fontface="bold")
```

